Incidence of confirmed COVID-19 cases in care home staff and residents: Cohort study using electronic records in a large UK care home provider (March-June 2020)

### 2020-06-30 covidWarehouse/manuscript/manuscript.Rmd
update data till 14 June

date_min <- as.Date("2020-03-02")
date_max <- as.Date("2020-06-07") #as.Date("2020-06-15")
date_min_tallies = as.Date("2020-03-24")
date_max_tallies = as.Date("2020-06-07")

@@

occupancy_interpolated <- resident_days_linear_approx() %>% 
  # TODO: currently putting the individual-level occup for 360 because no occ tally is known!
  # it's got 57 bed including 5 contract beds, so 
  left_join(transmute(filter(approx_occupancy, home_code == "360"), 
                      date, home_code  , occup360 = trunc(occupancy * 57/52)),
            by = c("home_code", "date")) %>%
  mutate(occupancy = if_else(home_code == "360", occup360, occupancy)) %>% 
  select(-occup360)

@@

reference_homes <- reference_homes %>% 
  left_join(home_mean_occupancy) %>% 
  mutate(pct_occupancy = mean_occupancy/beds)

residents_episodes <- tline %>% 
  filter(rday==1) %>% 
  group_by(resident_id, home_code, dob, gender, 
           status, home_type, user_type2,
           susceptible_symptomatic, 
           susceptible_confirmed) %>% 
  summarise(entry_date = min(date),
            exit_date = max(date+1),
            resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed)) %>% 
  mutate(date1 = as.numeric( entry_date - date_min),
         date2 = as.numeric(exit_date - date_min),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1),
         date2 = if_else(date1 == date2, date2 + 1, date2)) %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed")),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95))   %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category, beds, pct_occupancy)) %>% 
  group_by(resident_id) %>% 
  mutate(death_status = if_else(
    date1 == max(date1),
    as.integer(status == "Deceased"),
    0L
  )) %>% 
  ungroup()

residents_overview <- tline %>% 
  filter(rday==1) %>%
  mutate(ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  group_by(resident_id, home_code, ageg, gender, 
           admission_type2, home_type, user_type2, status) %>% 
  summarise(resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed),
            symptomatic = max(-susceptible_symptomatic+1, na.rm = T),
            confirmed = max(-susceptible_confirmed+1, na.rm = T),
            start_date = min(date, na.rm = T),
            stop_date = max(date, na.rm = T),
            start_int = as.numeric(min(date, na.rm = T) - date_min),
            stop_int = as.numeric(max(date, na.rm = T)- date_min)
    ) %>% 
  ungroup() %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed"))) %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category, beds, beds, pct_occupancy)) %>% 
  left_join(distinct(select(incidents, resident_id, starts_with("covid_first"))))

@@

home_total_confirmed <- distinct(reference_homes, home_code, home_name_clickview, postcode) %>% 
  left_join(select(reference_geography, postcode, country, region_UK), by = "postcode") %>% 
  left_join(res_counts, by = "home_code") %>% 
  left_join(home_total_confirmed_CT) %>% 
  left_join(home_total_confirmed_DAT) %>% 
  mutate(CT_confirmed = tidyr::replace_na(CT_confirmed, 0),
        DAT_confirmed = tidyr::replace_na(DAT_confirmed, 0))

@@

residents_ct <- tibble(date = seq(date_min, date_max_tallies, 1)) %>% 
  left_join(distinct(new_cases, date, home_code, 
                     new_confirmed_home, new_confirmed_hospital,
                     new_symptomatic_home, new_symptomatic_hospital)) %>% 
  filter(!is.na(home_code)) %>% 
  left_join(distinct(reference_homes, home_code, home_name_datix, postcode, staff_total, nursing_dementia, nursing_general, residential_dementia, residential_general, younger_person)) %>% 
  left_join(select(reference_geography, postcode, region_UK)) %>% 
  left_join(occupancy_interpolated) %>% 
  arrange(home_code, date) %>% 
  group_by(home_code) %>% 
  mutate(new_symptomatic = tidyr::replace_na(new_symptomatic_home, 0) +
                                    tidyr::replace_na(new_symptomatic_hospital, 0),
         new_confirmed = tidyr::replace_na(new_confirmed_home, 0) +
                                    tidyr::replace_na(new_confirmed_hospital, 0)) %>% 
  mutate(I_symptomatic = total_cases_in_home(occupancy, new_symptomatic, date),
         I_confirmed = total_cases_in_home(occupancy, new_confirmed, date)) %>% 
  ungroup()

km_ct_residents_confirmed <- residents_ct %>% 
  group_by(date, region_UK) %>% 
  summarise(
    confirmed = sum(new_confirmed),
    I_confirmed = round(sum(I_confirmed)),
    occupancy = sum(occupancy)
  ) %>% 
  group_by(region_UK) %>% 
  km_ct_estimator(df = ., t = date, d = confirmed, 
                  pop = (occupancy - I_confirmed), subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog)
  )

# residents_ct_symp <- tibble(date = seq(date_min, date_max_tallies, 1)) %>% 
#   left_join(distinct(new_cases, date, home_code, 
#                      new_symptomatic_home, new_symptomatic_hospital)) %>% 
#   filter(!is.na(home_code)) %>% 
#   left_join(distinct(reference_homes, home_code, home_name_datix, postcode)) %>% 
#   left_join(select(reference_geography, postcode, region_UK)) %>% 
#   left_join(occupancy_interpolated) %>% 
#   arrange(date)

km_ct_residents_symp <- residents_ct %>% 
  group_by(date, region_UK) %>% 
  summarise(
    symptomatic = sum(new_symptomatic),
    I_symptomatic = round(sum(I_symptomatic)),
    occupancy = sum(occupancy)
  ) %>% 
  group_by(region_UK) %>% 
  km_ct_estimator(df = ., t = date, d = symptomatic, 
                  pop = (occupancy - I_symptomatic), subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog)
  )

@@

km_staff_symptomatic <- staff_ct %>% 
  group_by(date, region_UK) %>% 
  summarise(
    symptomatic = sum(new_colleague_symptomatic),
    I_symptomatic = round(sum(I_symptomatic)),
    staff_total = sum(staff_total)
  ) %>% 
  group_by(region_UK) %>% 
  km_ct_estimator(., t=date, d=symptomatic, pop=(staff_total-I_symptomatic),
                  subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog)
  )

km_staff_confirmed <- staff_ct %>% 
  group_by(date, region_UK) %>% 
  summarise(
    confirmed = sum(new_colleague_confirmed),
    I_confirmed = round(sum(I_confirmed)),
    staff_total = sum(staff_total)
  ) %>% 
  group_by(region_UK) %>% 
  km_ct_estimator(., t=date, d=confirmed, pop=(staff_total-I_confirmed),
                  subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog)
  )

@@

```{r descriptives}
desc_residents <- list()
desc_residents$DAT_total_residents <- dplyr::n_distinct(residents_overview$resident_id)
desc_residents$DAT_total_residents_england <- dplyr::filter(residents_overview, country == "England") %>% summarise(n = dplyr::n_distinct(resident_id)) %>% .$n

#COUNTS
# This is estimated in validation.Rmd
desc_residents$CT_total_residents <- round(desc_residents$DAT_total_residents * 1.117309 + .49)
desc_residents$CT_total_residents_england <- round(desc_residents$DAT_total_residents_england * 1.0793268 + .49)
desc_residents$CT_total_symptomatic <- sum(new_cases$new_symptomatic_home, na.rm=T) + 
  sum(new_cases$new_symptomatic_hospital, na.rm=T)
desc_residents$CT_total_symptomatic_denom <- sum(km_ct_residents_symp$n_t)
desc_residents[paste0(
  "CT_p_symptomatic", c("", "_lower", "_upper")
)] <- epitools::binom.exact(desc_residents$CT_total_symptomatic,
                            desc_residents$CT_total_residents)[
                              c("proportion", "lower", "upper")
                            ] * 100

@@

desc_residents_outbreak_only$CT_total_residents <- residents_overview %>% 
  dplyr::filter(home_code %in% desc_residents_outbreak_only$outbreak_homes) %>% 
  dplyr::summarise(total_residents = round(dplyr::n_distinct(resident_id)* 1.1151324 + .49)) %>% 
  .$total_residents

@@

fig_KM_CT_residents_symptomatic <- fig_nums(
  name = "fig_KM_CT_residents_symptomatic",
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of residents becoming symptomatic based on care home manager daily counts")
)
fig_KM_residents_confirmed <- fig_nums(
  name = 'fig_KM_residents_confirmed',
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of residents (n=",
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ") becoming confirmed cases based on Datix incidents report"))
fig_KM_CT_confirmed_residents <- fig_nums(
  name = 'fig_KM_CT_confirmed_residents',
  caption = "Kaplan-Meier estimates of the cumulative probability of residents becoming confirmed cases based on care home manager daily counts")
fig_KM_mortality <- fig_nums(
  "fig_KM_mortality",
  caption = paste0("Kaplan-Meier estimates of resident (n=",
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ") survival by SARS-COV-2 case type"))
fig_KM_staff_symptomatic <- fig_nums(
  name = 'fig_KM_staff_symptomatic',
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of care home staff (n=",   formatbm(desc_staff$total_staff),
  ") becoming symptomatic cases based on manager-reported daily counts"))
fig_KM_staff_confirmed <- fig_nums(
  name = 'fig_KM_staff_confirmed',
  caption = paste0("Kaplan-Meier estimates of the cumulative probability of care home staff (n=",   formatbm(desc_staff$total_staff),
  ") becoming confirmed cases based on manager-reported daily counts"))

@@

```{r period_prevalence}
cis_data <- list(
  n_positive = 66,
  n_negative = 60618
)

england_results <- list()
england_results$n_positive = sum(inner_join(new_cases, filter(reference_homes, country == "England"))[c("new_confirmed_home", "new_confirmed_hospital")], na.rm = T)
england_results$n_negative = desc_residents$CT_total_residents_england  - england_results$n_positive

england_results$RR_england <- relative_risk(
  a = england_results$n_positive,
  b = england_results$n_negative,
  c = cis_data$n_positive,
  d = cis_data$n_negative)

england_results$RR_SE <- relative_risk_SE(
  a = england_results$n_positive,
  b = england_results$n_negative,
  c = cis_data$n_positive,
  d = cis_data$n_negative)
```

@@

**Findings:** An estimated `r formatbm(desc_residents$CT_total_residents)` residents were present in `r dplyr::n_distinct(residents$home_code)` FSHCG homes between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")`. During the period `r formatbm(desc_residents$CT_total_symptomatic_denom)` residents developed COVID-19 symptoms (`r tbrounding(desc_residents$CT_p_symptomatic)`% [95% confidence interval: `r tbrounding(desc_residents$CT_p_symptomatic_lower)`%; `r tbrounding(desc_residents$CT_p_symptomatic_upper)`%]), and `r formatbm(desc_residents$CT_total_confirmed)` (`r tbrounding(desc_residents$CT_p_confirmed)`% [`r tbrounding(desc_residents$CT_p_confirmed_lower)`%; `r tbrounding(desc_residents$CT_p_confirmed_upper)`%]) had an infection confirmed with nasopharyngeal RNA tests. Across all care homes, the confirmed infection rate of incidence was `r tbrounding(desc_residents$CT_IR_confirmed)` per 100,000 resident-days [`r tbrounding(desc_residents$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents$CT_IR_confirmed_upper)`]. In England, `r england_results$n_positive` infections were confirmed of a total `r formatbm(england_results$n_positive + england_results$n_negative)` residents. Comparing with the community, the relative risk of confirmed cases in England was `r tbrounding(england_results$RR_england)` [`r tbrounding(exp(log(england_results$RR_england) - 1.96 * england_results$RR_SE)) `; `r tbrounding(exp(log(england_results$RR_england) + 1.96 * england_results$RR_SE))`]. These estimates include `r desc_residents$CT_zero_case`/`r desc_residents$total_homes` homes (`r tbrounding( desc_residents$CT_zero_case/desc_residents$total_homes*100)`%) that had not reported a single confirmed case.

The staff-to-bed ratio was the only factor to be independently associated with the rate of confirmed cases: an increase by 10 percentage point in the number of staff relative to the bed total was associated with a 13% reduction [9%, 18%] in the rate of confirmed cases. Out of `r formatbm(desc_staff$total_staff)` staff, `r formatbm(desc_staff$CT_confirmed_total)` confirmed infections were counted(
`r tbrounding(desc_staff$CT_confirmed_p)`% [`r tbrounding(desc_staff$CT_confirmed_p_lower)`%; `r tbrounding(desc_staff$CT_confirmed_p_upper)`%]), corresponding to a rate of incidence of 
`r tbrounding(desc_staff$CT_confirmed_IR)` per 100,000 staff-days [`r tbrounding(desc_staff$CT_symptomatic_IR_lower)`; `r tbrounding(desc_staff$CT_symptomatic_IR_upper)`].

@@

## Aggregate outbreak data (`r format(date_min_tallies, "%d/%m/%Y")`--`r format(date_max_tallies, "%d/%m/%Y")` )

@@

## Individual-level residents data (`r format(date_min, "%d/%m/%Y")`--`r format(date_max, "%d/%m/%Y")` )

@@

Incidence proportions and rates were calculated using the aggregate outbreak counts, since they were the trusted source of information and encompass all residents. The number of unique residents $N$ present in the homes at some point between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")`  was estimated by multiplying the number of private residents (from the individual-level dataset) by the ratio of the total resident-days (from the weekly occupancy censuses) to the total private resident-days (from the individual data). This approach assumes that the mean stay of local authority residents is identical to that of private residents (supplementary material 1). As the total exposed-to-risk (susceptible resident-days) was unknown, it was approximated in a multiple decrement life table by making the assumption that previously symptomatic/confirmed residents were discharged at the same rate as susceptible residents, as was suggested by the individual-level records. Admitted cases were accounted for in the daily new case count. Based on this $\hat{E_t}$, the estimated number or previously symptomatic/confirmed residents present in the home on day $t$ was computed as

>>>>

date_min <- as.Date("2020-03-02")
date_max <- as.Date("2020-06-14")
date_min_tallies = as.Date("2020-03-24")
date_max_tallies = as.Date("2020-06-14")

@@

occupancy_interpolated <- resident_days_linear_approx() #%>% 
  # # TODO: currently putting the individual-level occup for 360 because no occ tally is known!
  # # it's got 57 bed including 5 contract beds, so 
  # left_join(transmute(filter(approx_occupancy, home_code == "360"), 
  #                     date, home_code  , occup360 = trunc(occupancy * 57/52)),
  #           by = c("home_code", "date")) %>%
  # mutate(occupancy = if_else(home_code == "360", occup360, occupancy)) %>% 
  # select(-occup360)

@@

reference_homes <- reference_homes %>% 
  left_join(home_mean_occupancy) %>% 
  mutate(pct_occupancy = mean_occupancy/beds,
         occ_bedroom_ratio = mean_occupancy/no_bedrooms)

residents_episodes <- tline %>% 
  filter(rday==1) %>% 
  group_by(resident_id, home_code, dob, gender, 
           status, home_type, user_type2,
           susceptible_symptomatic, 
           susceptible_confirmed) %>% 
  summarise(entry_date = min(date),
            exit_date = max(date+1),
            resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed)) %>% 
  mutate(date1 = as.numeric( entry_date - date_min),
         date2 = as.numeric(exit_date - date_min),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1),
         date2 = if_else(date1 == date2, date2 + 1, date2)) %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed")),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95))   %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category, 
                   beds, pct_occupancy, occ_bedroom_ratio, no_bedrooms)) %>% 
  group_by(resident_id) %>% 
  mutate(death_status = if_else(
    date1 == max(date1),
    as.integer(status == "Deceased"),
    0L
  )) %>% 
  ungroup()

residents_overview <- tline %>% 
  filter(rday==1) %>%
  mutate(ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  group_by(resident_id, home_code, ageg, gender, 
           admission_type2, home_type, user_type2, status) %>% 
  summarise(resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed),
            symptomatic = max(-susceptible_symptomatic+1, na.rm = T),
            confirmed = max(-susceptible_confirmed+1, na.rm = T),
            start_date = min(date, na.rm = T),
            stop_date = max(date, na.rm = T),
            start_int = as.numeric(min(date, na.rm = T) - date_min),
            stop_int = as.numeric(max(date, na.rm = T)- date_min)
    ) %>% 
  ungroup() %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed"))) %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category,
                   beds,  pct_occupancy, occ_bedroom_ratio, no_bedrooms)) %>% 
  left_join(distinct(select(incidents, resident_id, starts_with("covid_first"))))

@@

home_total_confirmed <- distinct(reference_homes, home_code, home_name_clickview, 
                                 country, region_UK, staff_to_bed_ratio, carer_to_bed_ratio, 
                                 bed_tot_category, beds, pct_occupancy, no_bedrooms, occ_bedroom_ratio,
                                 younger_person, residential_general, residential_dementia,
                                 nursing_general,nursing_dementia) %>% 
  left_join(res_counts, by = "home_code") %>% 
  left_join(home_total_confirmed_CT) %>% 
  left_join(home_total_confirmed_DAT) %>% 
  mutate(CT_confirmed = tidyr::replace_na(CT_confirmed, 0),
         DAT_confirmed = tidyr::replace_na(DAT_confirmed, 0))

@@

residents_ct <- tibble(date = seq(date_min, date_max_tallies, 1)) %>% 
  left_join(select(new_cases, date, home_code, 
                     new_confirmed_home, new_confirmed_hospital,
                     new_symptomatic_home, new_symptomatic_hospital,
                     new_deaths_home, new_deaths_hospital)) %>% 
  filter(!is.na(home_code)) %>% 
  left_join(distinct(reference_homes, home_code, home_name_datix, postcode, 
                     staff_total, nursing_dementia, nursing_general, 
                     residential_dementia, residential_general, younger_person)) %>% 
  left_join(select(reference_geography, postcode, region_UK)) %>% 
  left_join(occupancy_interpolated) %>% 
  arrange(home_code, date) %>% 
  group_by(home_code) %>% 
  mutate(new_symptomatic = tidyr::replace_na(new_symptomatic_home, 0) +
                                    tidyr::replace_na(new_symptomatic_hospital, 0),
         new_confirmed = tidyr::replace_na(new_confirmed_home, 0) +
                                    tidyr::replace_na(new_confirmed_hospital, 0),
         new_deaths = tidyr::replace_na(new_deaths_home, 0) +
                                    tidyr::replace_na(new_deaths_hospital, 0)) %>% 
  mutate(I_symptomatic = total_cases_in_home(occupancy, new_symptomatic, date),
         I_confirmed = total_cases_in_home(occupancy, new_confirmed, date)) %>% 
  ungroup()

km_ct_residents_deaths <- residents_ct %>% 
  group_by(date) %>% 
  summarise(
    deaths = sum(new_deaths),
    occupancy = sum(occupancy)
  ) %>% 
  km_ct_estimator(df = ., t = date, d = deaths, 
                  pop = occupancy, subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog),
    event = "Death"
  )

km_ct_residents_confirmed <- residents_ct %>% 
  group_by(date) %>% 
  summarise(
    confirmed = sum(new_confirmed),
    I_confirmed = round(sum(I_confirmed)),
    occupancy = sum(occupancy)
  ) %>% 
  km_ct_estimator(df = ., t = date, d = confirmed, 
                  pop = (occupancy - I_confirmed), subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog),
    event = "Confirmed infection"
  )

km_ct_residents_symp <- residents_ct %>% 
  group_by(date) %>% 
  summarise(
    symptomatic = sum(new_symptomatic),
    I_symptomatic = round(sum(I_symptomatic)),
    occupancy = sum(occupancy)
  ) %>% 
  km_ct_estimator(df = ., t = date, d = symptomatic, 
                  pop = (occupancy - I_symptomatic), subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog),
    event = "Symptomatic case"
  )

@@

km_staff_symptomatic <- staff_ct %>% 
  group_by(date) %>% 
  summarise(
    symptomatic = sum(new_colleague_symptomatic),
    I_symptomatic = round(sum(I_symptomatic)),
    staff_total = sum(staff_total)
  ) %>% 
  km_ct_estimator(., t=date, d=symptomatic, pop=(staff_total-I_symptomatic),
                  subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog),
    event = "Symptomatic case"
  )

km_staff_confirmed <- staff_ct %>% 
  group_by(date) %>% 
  summarise(
    confirmed = sum(new_colleague_confirmed),
    I_confirmed = round(sum(I_confirmed)),
    staff_total = sum(staff_total)
  ) %>% 
  km_ct_estimator(., t=date, d=confirmed, pop=(staff_total-I_confirmed),
                  subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog),
    event = "Confirmed infection"
  )

@@

```{r descriptives}
desc_residents <- list()
desc_residents$DAT_total_residents <- dplyr::n_distinct(residents_overview$resident_id)

#COUNTS
# This is estimated in validation.Rmd
desc_residents$CT_total_residents <- tline %>% 
  dplyr::filter(between(date, date_min_tallies, date_max_tallies) & rday == 1) %>% summarise(n = round(dplyr::n_distinct(resident_id) * 1.123865 + .49)) %>% .$n

desc_residents$CT_total_residents_england <- tline %>% 
  dplyr::inner_join(select(filter(reference_homes, country == "England"), home_code)) %>%
  dplyr::filter(between(date, date_min_tallies, date_max_tallies) & rday == 1) %>%
  summarise(n = round(dplyr::n_distinct(resident_id) * 1.087621 + .49)) %>% .$n

desc_residents$CT_total_symptomatic <- sum(new_cases$new_symptomatic_home, na.rm=T) + 
  sum(new_cases$new_symptomatic_hospital, na.rm=T)
desc_residents$CT_total_symptomatic_denom <- sum(km_ct_residents_symp$n_t)
desc_residents[paste0(
  "CT_p_symptomatic", c("", "_lower", "_upper")
)] <- epitools::binom.exact(desc_residents$CT_total_symptomatic,
                            desc_residents$CT_total_residents)[
                              c("proportion", "lower", "upper")
                            ] * 100

@@

desc_residents_outbreak_only$CT_total_residents <- residents_overview %>% 
  dplyr::filter(home_code %in% desc_residents_outbreak_only$outbreak_homes) %>% 
  dplyr::summarise(total_residents = round(dplyr::n_distinct(resident_id)* 1.122205 + .49)) %>% 
  .$total_residents

@@

fig_KM_CT_residents <- fig_nums(
  name = "fig_KM_CT_residents",
  caption = paste0("Kaplan-Meier estimates of the cumulative incidence of symptomatic cases, confirmed infections and COVID-related deaths in residents according to FSHCG outbreak surveillance reports (", format(date_min_tallies, "%d/%m/%y"), "-", format(date_max_tallies, "%d/%m/%y"), ")")
)

fig_KM_staff <- fig_nums(
  name = 'fig_KM_staff',
  caption = paste0("Kaplan-Meier estimates of the cumulative incidence of symptomatic cases, confirmed infection in care home staff (n=",   formatbm(desc_staff$total_staff),
  ") according to FSHCG outbreak surveillance reports (", format(date_min_tallies, "%d/%m/%y"), "-", format(date_max_tallies, "%d/%m/%y"), ")"))

fig_KM_mortality <- fig_nums(
  "fig_KM_mortality",
  caption = paste0("Kaplan-Meier estimates of resident (n=",
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ") survival by SARS-COV-2 case type"))

@@

```{r period_prevalence, message=F}
# Cumulative sample counts, directly standardised for
# 11 May to 24 May
# 25 May to 7 June

england_results <- list(
  date_min = as.Date("2020-05-11"),
  date_max = as.Date("2020-06-07"),
  cis_confirmed = 35,
  cis_person_days = 483259
)

england_results$FSHC_confirmed <- new_cases %>% 
  filter(between(date, england_results$date_min, england_results$date_max)) %>% 
  inner_join(filter(reference_homes, country == "England")) %>% 
  summarise(
    n = sum(new_confirmed_home, na.rm = T) + sum(new_confirmed_hospital, na.rm = T)
  ) %>% 
  .$n

england_results$FSHC_resident_days <- filter(km_ct_residents_confirmed,
  between(date, england_results$date_min, england_results$date_max)) %>% 
  summarise(n = sum(n_t, na.rm = T) ) %>% 
  .$n

england_results$results <- epitools::rateratio.wald(
  c(england_results$cis_confirmed,
    england_results$FSHC_confirmed,
    england_results$cis_person_days,
    england_results$FSHC_resident_days))
```

@@

**Findings:** An estimated `r formatbm(desc_residents$CT_total_residents)` residents were present in `r dplyr::n_distinct(residents$home_code)` FSHCG homes between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")`. During the period `r formatbm(desc_residents$CT_total_symptomatic_denom)` residents developed COVID-19 symptoms (`r tbrounding(desc_residents$CT_p_symptomatic)`% [95% confidence interval: `r tbrounding(desc_residents$CT_p_symptomatic_lower)`%; `r tbrounding(desc_residents$CT_p_symptomatic_upper)`%]), and `r formatbm(desc_residents$CT_total_confirmed)` (`r tbrounding(desc_residents$CT_p_confirmed)`% [`r tbrounding(desc_residents$CT_p_confirmed_lower)`%; `r tbrounding(desc_residents$CT_p_confirmed_upper)`%]) had an infection confirmed with nasopharyngeal RNA tests. Across all care homes, the confirmed infection rate of incidence was `r tbrounding(desc_residents$CT_IR_confirmed)` per 100,000 resident-days [`r tbrounding(desc_residents$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents$CT_IR_confirmed_upper)`].  These estimates include `r desc_residents$CT_zero_case`/`r desc_residents$total_homes` homes (`r tbrounding( desc_residents$CT_zero_case/desc_residents$total_homes*100)`%) that had not reported a single confirmed case.
In a comparison with results from an English household survey, we estimated a confirmed infection rate ratio of `r tbrounding(england_results$results$measure[2,"estimate"])` [`r tbrounding(england_results$results$measure[2,"lower"]) `; `r tbrounding(england_results$results$measure[2,"upper"])`] in care homes compared with the community in England between `r format(england_results$date_min, "%e %B %Y")` and `r format(england_results$date_max, "%e %B %Y")`.
The staff-to-bed ratio was the main organisational factor to be independently associated with the rate of confirmed cases: an increase by 10 percentage point in the number of staff relative to the bed total was associated with a 13% reduction [9%, 18%] in the rate of confirmed cases. Out of `r formatbm(desc_staff$total_staff)` staff, `r formatbm(desc_staff$CT_confirmed_total)` confirmed infections were counted(
`r tbrounding(desc_staff$CT_confirmed_p)`% [`r tbrounding(desc_staff$CT_confirmed_p_lower)`%; `r tbrounding(desc_staff$CT_confirmed_p_upper)`%]), corresponding to a rate of incidence of 
`r tbrounding(desc_staff$CT_confirmed_IR)` per 100,000 staff-days [`r tbrounding(desc_staff$CT_symptomatic_IR_lower)`; `r tbrounding(desc_staff$CT_symptomatic_IR_upper)`].

@@

## Aggregate outbreak data (`r format(date_min_tallies, "%d/%m/%Y")`--`r format(date_max_tallies, "%d/%m/%Y")`)

@@

## Individual-level residents data (`r format(date_min, "%d/%m/%Y")`--`r format(date_max, "%d/%m/%Y")`)

@@

Incidence proportions and rates were calculated using the aggregate outbreak counts, since they were the trusted source of information and encompass all residents. The number of unique residents $N$ present in the homes at some point between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")`  was estimated by multiplying the number of private residents (from the individual-level dataset) by the ratio of the total resident-days (from the weekly occupancy censuses) to the total private resident-days (from the individual data). This approach assumes that the mean stay of local authority residents is identical to that of private residents (supplementary material 1). As the total exposed-to-risk (susceptible resident-days) was unknown, it was approximated in a multiple decrement life table by making the assumption that previously symptomatic/confirmed residents were discharged at the same rate as susceptible residents, as was suggested by the individual-level records. Admitted cases were accounted for in the daily new case count. Based on this $\hat{I_t}$, the estimated number or previously symptomatic/confirmed residents present in the home on day $t$ was computed as

$$\hat{I_t} = min(n_{t}/n_{t-1}, 1)( \hat{E}_{t-1} + d_{t-1}); \qquad  \qquad  \qquad I_{t_0} = 0$$
where $d_{t}$ denotes the number of new cases for each day $t = \{0, \ldots, T\}$, while $min(n_{t}/n_{t-1}, 1)$ acts as an (underestimated) approximation of the discharge rate. The total exposure time was calculated as $\sum_{t=0}^{T} (N_t-\hat{I}_t)$ where $N_t$ denotes daily occupancy figures. The incidence proportion of confirmed cases in England was contrasted with incidence measured from a sample survey of English private households which tested nose/throat swabs from 26,995 individuals between 26 April and 13 June 2020.[@cisbulletin2020; @cisdata2020]
